
#Testing loading a package
```{r, include=FALSE}
library(ggplot2)
library(lme4)
library(nlme)
library(car)
library(lsmeans)
library(lmerTest)
```
##Statistical analyses
###Wild, urban population  
For each bird, calculated the condition as a regression of mass and tarsus length, using the following formula: 
`Condit<-lm(Mass~Tarsus)`
```{r, echo=FALSE}
fwild<-read.csv(file="~/Documents/side projects/Wild_feeder cleaning_analysis_corrected.csv")
Condit <- lm(Mass ~ Tarsus, data=fwild, na.action=na.exclude)
#significant and positive - good
summary(Condit)
#put residuals in dataframe
fwild$Condition <- residuals(Condit)
```
Because some wild birds were recaptures and sampled more than once, I then used a random number generator in R to create a random ID and select one data point, randomly, from birds with more than one data point during the study.
This then changed the sample size between the weeks. The table below shows the week # (first row) with sample size below. 
```{r, echo=FALSE} 
#UNIQUE ID SORTING WITH RANDOM NUMBERS
#to analyse with only one measurement per bird, randomly chosen...
#HAVE ALREADY CALCULATED CONDITION BEFORE THIS STEP!

#THIS STEP WILL RE-DO EVERY TIME IF ALLOWED TO, CHANGING THE DATASET. uniqWild dataset should be made permanent at some point. 
#random number generator - to sort and select only one reading from recapped birds.
fwild$randomID <- sample(100, size = nrow(fwild), replace = TRUE)
#sort samples by random ID
uniqWild<-fwild[with(fwild, order(ID, randomID)), ]
#get rid of duplicate IDs, keeping only the first ID in random order
uniqWild<-uniqWild[ !duplicated(uniqWild$ID), ]
#count sample size per week
table(uniqWild$Week)
```
```{r, echo=FALSE}
#USING CYCLE TIME instead of WEEK
fwild$time<-0
fwild$time[fwild$Week==0]<-"NA"
fwild$time[fwild$Week==1]<-"1"
fwild$time[fwild$Week==2]<-"1"
fwild$time[fwild$Week==3]<-"1"
fwild$time[fwild$Week==4]<-"1"
fwild$time[fwild$Week==5]<-"2"
fwild$time[fwild$Week==6]<-"2"
fwild$time[fwild$Week==7]<-"2"
fwild$time[fwild$Week==8]<-"2"

uniqWild$time<-0
uniqWild$time[uniqWild$Week==0]<-"0"
uniqWild$time[uniqWild$Week==1]<-"1"
uniqWild$time[uniqWild$Week==1]<-"1"
uniqWild$time[uniqWild$Week==3]<-"1"
uniqWild$time[uniqWild$Week==4]<-"1"
uniqWild$time[uniqWild$Week==5]<-"2"
uniqWild$time[uniqWild$Week==6]<-"2"
uniqWild$time[uniqWild$Week==7]<-"2"
uniqWild$time[uniqWild$Week==8]<-"2"

#write updated fwild datatable to a file
write.csv(fwild,file="~/Documents/side projects/wild_feeder_study_edited.csv")
```
```{r, echo=FALSE}
#GRAPHS
#The following code summarizes the data, using a custom function, and generates the values needed for confidence intervals, standard errors, etc. used in graph functions. 
#summarize data function (defines the function)
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}
#running data summary function
fwild.data <- data_summary(fwild, varname="Condition", 
                    groupnames=c("FeederStatus", "Week"))
uniqWild.data <- data_summary(uniqWild, varname="Condition", 
                    groupnames=c("FeederStatus", "Week"))
# Convert Week to a factor variable
fwild.data$Week=as.factor(fwild.data$Week)
uniqWild.data$Week=as.factor(uniqWild.data$Week)
#head(fwild.data)  #uncomment this to verify the function is working
fwild.data2 <-data_summary(fwild, varname="Coccidia",
                           groupnames=c("FeederStatus", "time"))
uniqWild.data2 <-data_summary(uniqWild, varname="Coccidia",
                           groupnames=c("FeederStatus", "time"))
#import summarySE formula from here: summarySE.R
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
fwild.ci<-summarySE(fwild, measurevar="Condition", groupvars=c("Week", "FeederStatus"), na.rm=TRUE) #Condition
fwild.ci2<-summarySE(fwild, measurevar="Coccidia", groupvars=c("Week", "FeederStatus"), na.rm=TRUE) #Coccidians
uniqWild.ci<-summarySE(uniqWild, measurevar="Condition", groupvars=c("Week", "FeederStatus"), na.rm=TRUE) #Condition
uniqWild.ci2<-summarySE(uniqWild, measurevar="Coccidia", groupvars=c("Week", "FeederStatus"), na.rm=TRUE) #Coccidians
```
The below graphs show the trend of body condition in the wild-caught birds over the course of the study. 
```{r, echo=FALSE}
#ggplot GRAPHS
library(ggplot2)
#boxplot (median, variance) Condition by FeederStatus
fwild$Week<-as.factor(fwild$Week)
p3<-ggplot(fwild, aes(x=Week, y=Condition, fill=FeederStatus)) + geom_boxplot() +
  ggtitle("Body condition by week (wild, urban birds)", subtitle = "Boxplots: central lines represent median")
print(p3)

# uniqWild$Week<-as.factor(uniqWild$Week)
# p3a<-ggplot(uniqWild, aes(x=Week, y=Condition, fill=FeederStatus)) + geom_boxplot() +
#   ggtitle("Body condition by week [Unique wild birds]", subtitle = "Boxplots: central lines represent median ")
# print(p3a)

pd <- position_dodge(0.1) 
#condition with confidence intervals
p4<-ggplot(fwild.ci, aes(x=Week, y=Condition, colour=FeederStatus)) + 
  geom_errorbar(aes(ymin=Condition-ci, ymax=Condition+ci), width=.2,
                position=pd) +
  geom_point(position=pd) +
  ggtitle("Body condition by week (wild, urban birds)", subtitle = "Means, error bars represent 95% confidence intervals")
print(p4)

# p4a<-ggplot(uniqWild.ci, aes(x=Week, y=Condition, colour=FeederStatus)) + 
#   geom_errorbar(aes(ymin=Condition-ci, ymax=Condition+ci), width=.2,
#                 position=pd) +
#   geom_point(position=pd) +
#   ggtitle("Body condition by week [Unique wild birds]", subtitle = "Means, error bars represent 95% confidence intervals")
# print(p4a)
```
A type-II ANOVA, using unique bird IDs only (no recapture data included), shows a significant effect of week on body condition, but _not_ feeder status. Additionally, there is no significant interaction between feeder status and week, suggesting the effect is of time, but not the feeder treatment "cycle" number. 
```{r, echo=FALSE}
Anova(lm(Condition ~ FeederStatus*Week, data = uniqWild))
Anova(lmer(Condition ~ FeederStatus*Week + (1|ID) +(1|time), data=fwild))
```
Using likelihood tests of mixed models, time (cycle number 1 or 2) significantly affects body condition, compared to the reduced null model. However, incorporating feeder status as a fixed effect does not significantly increase likelihood. Additionally, no interaction between feeder status and time is significant. 
```{r,echo=FALSE}
wild.red<- lmer(Condition ~ 1 + (1|ID) + (1|Week) , data=fwild, REML=FALSE)
wild.null<-lmer(Condition ~ time + (1|ID) + (1|Week), data=fwild,REML=FALSE)
wild1<-lmer(Condition ~ FeederStatus + time+ (1|ID) + (1|Week), data=fwild,REML=FALSE)
wild2<-lmer(Condition ~ FeederStatus*time+ (1|ID) + (1|Week), data=fwild,REML=FALSE)
anova(wild.red,wild.null)
anova(wild1, wild.null)
anova(wild2, wild1)
```
Next are plots of coccidia occurence in the wild,urban population across the study. 
Coccidia was collected from fecal samples and was scored from 0 - 4. 
```{r,echo=FALSE}
#coccidia boxplot
p5<-ggplot(fwild, aes(x=Week, y=Coccidia, fill=FeederStatus)) + geom_boxplot()+
  ggtitle("Coccidia by Week (wild, urban birds)", subtitle = "Boxplots with medians")
print(p5)
#coccidia line graph
p6<-ggplot(fwild.ci2, aes(x=Week, y=Coccidia, colour=FeederStatus)) + 
  geom_errorbar(aes(ymin=Coccidia-ci, ymax=Coccidia+ci), width=.2,
                position=pd) +
  geom_point(position=pd) +
  ylim(0,4) + 
  ggtitle("Coccidia by Week (wild, urban birds)", subtitle ="Means with 95% confidence intervals")
print(p6)
```
Using model likelihood ratio test, cycle number (time) showed a trend towards affecting coccidia load, however, this was not significant. Feeder status did not significantly affect coccidia load, and there was no significant interaction between feeder status and cycle number. 
```{r, echo=FALSE}
cocc.red<- lmer(Coccidia ~ 1 + (1|ID) + (1|Week), data=fwild, REML=FALSE)
cocc.null<- lmer(Coccidia ~ time + (1|ID) + (1|Week), data=fwild, REML=FALSE)
cocc1<- lmer(Coccidia ~ FeederStatus + time + (1|ID) + (1|Week), data=fwild, REML=FALSE)
cocc2<- lmer(Coccidia ~ FeederStatus*time + (1|ID) + (1|Week), data=fwild, REML=FALSE)
anova(cocc.red, cocc.null)
anova(cocc.null, cocc1)
anova(cocc1, cocc2)
```
